Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
contribs

Experiments    with   a    single    generation   copying    collector
(Section~\ref{sec:experiments})  confirm   the  expected  performance
  benefits.   Liveness-based  collection  results in  an  increase  in
  garbage reclaimed.   As a  consequence, there is  a decrease  in the
  number of  collections (up to 45\%)  and a decrease (up  to 88\%) in
  the number  of {\em dragged}  cells (cells  that survive a  GC after
  their  last  use). Besides  there  is  a  reduction in  the  overall
  execution time for some programs.  Our experiments show that a mixed
  strategy, where we  do liveness based GC up to  the closure boundary
  and reachability based GC for  the arguments (free variables) of the
  closure,  gives us  similar benefits  as a  fully liveness  based GC
  (where  the  arguments  of  the closure  are  also  collected  using
  liveness) at much less overheads (Section~\ref{sec:strategies}).

@@

{\color {Myblue}Rewrite at the end: The specific contributions of the
paper
  are: The formulation of a  liveness analysis for a lazy language and
  the proof of its correctness, the undecidability of the grammar that
  results from the analysis and  its approximation by DFAs, the design
  of a  garbage collector that can  use the DFAs to  retain live cells
  that  also include  closures, and  empirical results  that  show the
  effectiveness of liveness-based garbage collection.}

>>>>

Experiments    with   a    single    generation   copying    collector
(Section~\ref{sec:experiments})   confirm  the   expected  performance
benefits.  Liveness-based collection results in an increase in garbage
reclaimed.  As  a consequence, there  is a  decrease in the  number of
collections (up to 45\%) and a decrease  (up to 88\%) in the number of
{\em  dragged}  cells  (cells  that  survive a  GC  after  their  last
use). Besides, there is a reduction  in the overall execution time for
some programs.  Our  experiments show that a mixed  strategy, where we
do liveness based GC up to the closure boundary and reachability based
GC for the arguments (free variables) of the closure, gives us similar
benefits as  a fully  liveness based  GC (where  the arguments  of the
closure  are  also  collected  using liveness)  with  lesser  overhead
(Section~\ref{sec:strategies}).

@@

The specific contributions of this work are
  \begin{itemize}
  \item  We formulate  a  liveness  analysis for  a  lazy first  order
    functional language. We show how  to compute function summaries in
    a  context  independent  manner  to achievs  a  context  sensitive
    analysis. We  further describe the transformation  of the liveness
    equations to grammar rules in a  way that the language accepted by
    the resulting grammar describes live heap locations.\mycomment{the
      proof of its correctness}
  \item We prove  the undecidability of the grammar  that results from
    the  analysis.   To overcome  this,  we  present an  algorithm  to
    transform grammar  to finite  state automata  (DFA) such  that the
    language accepted by the  DFAs over-approximates the liveness (and
    hence a safe approximation).
  \item We have implemented a garbage  collector that can use the DFAs
    to  retain live  cells that  also include  closures. We  have done
    extensive  experiments with  this garbage  collector and  obtained
    empirical results  that show  the effectiveness  of liveness-based
    garbage collection.
  \end{itemize}
